---
title: "Hayman Fire Recovery"
author: "Matthew Ross"
date: "10/3/2019"
output: html_document:
  theme: cerulean
  toc: true
      toc_float: true
      toc_depth: 3
editor_options: 
  chunk_output_type: console
---

```{r setup, warning=F,message=F}
library(tidyverse)
library(tidyr)
library(ggthemes)
library(lubridate)

# Now that we have learned how to munge (manipulate) data
# and plot it, we will work on using these skills in new ways

knitr::opts_knit$set(root.dir='..')
```


```{r dataread, warning=F,message=F}
####-----Reading in Data and Stacking it ----- ####
#Reading in files
files <- list.files("data", full.names=T)
# If run from line, lowest directory: need ../ | If run from console, base directory at .Rproj

paste(files)

#Read in individual data files
ndmi <- read_csv(files[1]) %>% 
  rename(burned=2,unburned=3) %>%
  mutate(data='ndmi')

ndsi <- read_csv(files[2]) %>% 
  rename(burned=2,unburned=3) %>%
  mutate(data='ndsi')

ndvi <- read_csv(files[3])%>% 
  rename(burned=2,unburned=3) %>%
  mutate(data='ndvi')

# Stack as a tidy dataset
full_long <- rbind(ndvi,ndmi,ndsi) %>%
  gather(key='site',value='value',-DateTime,-data) %>%
  filter(!is.na(value))


```




## Question 1) 

What is the correlation between NDVI and NDMI? - here I want you to
convert the full_long dataset in to a wide dataset using the 
function "spread" and then make a plot that shows the correlations a
function of if the site was burned or not (x axis should be ndmi)
You should exclude winter months and focus on summer months

```{r}
wide_long <- spread(full_long, key = "data", value = "value") %>% 
  mutate(month = month(DateTime)) %>% 
  filter(month %in% c(5:9))

ggplot(wide_long, aes(ndmi, ndvi, color = site)) + geom_point()
```



## Question 2 

2) What is the correlation between average NDSI (normalized
 snow index) for January - April and average NDVI for June-August?
In other words, does the previous year's snow cover influence vegetation
 growth for the following summer?

```{r}



wide_longearly <- spread(full_long, key = "data", value = "value") %>%
  mutate(month = month(DateTime)) %>%
  filter(month %in% c(1:4)) %>% 
  

wide_longlate <- spread(full_long, key = "data", value = "value") %>% 
  mutate(month = month(DateTime)) %>% 
  filter(month %in% c(6:8))



early <- wide_longearly %>% select(DateTime, ndsi) %>% 
  mutate(observation = 1:n())

late <- wide_long %>% select(DateTime, ndvi) %>% 
  mutate(observation = 1:n())

total <- full_join(early, late, by = "observation")

# total = merge(early$ndsi, late$ndvi) %>%
#   rename(ndsi = "x", ndvi = "y")

total_mod <- pivot_longer(total,cols = c(ndsi, ndvi), names_to = "type", values_to = "value")

ggplot(total_mod, aes(observation, value, color = type)) + geom_point()

totaledit <- total %>% na.omit()

cor(totaledit$ndsi, totaledit$ndvi)
```


## Q3

How is the snow effect from question 2 different between pre- and post-burn
and burned and unburned? 


## Question 4

What month is the greenest month on average? 




## Question 5) 

What month is the snowiest on average?


## Bonus Question: Redo all problems with `spread` and `gather` using modern tidyverse syntax. 


## Bonus Question: Use Climate Engine to pull the same data for the assignment, but updated with 2020/2021 data.




